Multistimulus Machine Learning in Microfluidic Memristor Networks

### 2025-06-06 paper_overleaf/paper.tex
Modified vd radius base.

\title{Efficient Machine Learning in a Memristor Network}

\author{Monica Conte}
\affiliation{Soft Condensed Matter \& Biophysics, Debye Institute for Nanomaterials Science, Utrecht University, Princetonplein 1, 3584 CC Utrecht, The Netherlands.}

@@

\begin{abstract}
Inspired by the brain's energy-efficient processing, we propose a physical implementation of artificial neural networks to surpass digital systems in speed and energy efficiency. We consider a network with nodes connected by conical microfluidic memristor channels filled with aqueous electrolyte [1]. These memristors exhibit tunable conductance, responsive to electrical, chemical, and mechanical stimuli, as well as channel geometry. Based on this adaptability, we develop a hybrid training algorithm that adjusts either channel geometry or external stimuli as weights. Physical relaxation replaces the feedforward pass, while a combined physical-numerical gradient descent drives the backward pass. This approach enables tasks such as input-output voltage mapping and linear regression with bias.
\end{abstract}
\maketitle

@@

Applications of PNNs have been explored across diverse fields, including optics \cite{optics1}, electronics \cite{electronics1}, photonics \cite{acoustic1} and mechanical systems \cite{mechanics1}, among others. Many existing training algorithms rely entirely on in silico methods, which often face challenges when transferring the learned model to real physical devices. To address this, hybrid training approaches have been proposed, where the inference phase is executed physically while the backward phase is computed numerically\cite{backporp_pnn}. Other methods, based on coupled learning\cite{supervised_learning}, implement local learning rules at the level of each connection or edge\cite{demonstr_coupled_learning}. Applying local rules removed the need of the optimiziation of a global function during training, as in standard backpropagation methods. However, these approaches still face the need for two identical physical systems to apply different boundary conditions. In this work, we propose a novel algorithm in which both inference and backpropagation are carried out mainly taking advantage of the physical processes inherent in the system, specifically leveraging the natural relaxation dynamics of the system in response to external stimuli and weight modifications.

@@

The paper is organized as follows. \cref{sec:memristor_network} presents the physical network we define using memristor elements, called memristor network, and explains how this system is numerically solved. \cref{sec:training} describes the training algorithm that we use to achieve desired responses of the memristor network. In \cref{sec:train_memr_vd}, we exemplify the proposed algorithm by training a simple system, namely a voltage divider, for input-output voltage mapping. \cref{sec:train_memr_nw} shows a bigger network able to be trained for more complex input-output relationships. Finally, in \cref{sec:linear_regression} we adapt the training algorthm to train a network to perform a linear regression with bias. This section is followed by conclusions and suggestions for future work in \cref{sec:concl}.

\newpage

@@

At the far side of both reservoirs, electrical potential, pressure and ion concentration are imposed. As a result, the memristor feels a potential drop $\Delta V = V_b - V_t$ defined as the difference between the potential in the base and in the tip. The ions within the EDL region generate an electro-osmotic fluid flow $Q_V = (\pi R_b R_t\epsilon\psi_0/L\eta)\Delta V$. Meanwhile, the pressure drop $\Delta P = P_b - P_t$, given by the difference between the base and the tip pressure, induces a Poiseuille-like flow $Q_P = (\pi R_b^3 R_t^3/8L\eta\langle R^2\rangle)\Delta P$. Finally, the ions experience a concentration gradient $\Delta \rho = \rho_b - \rho_t$, defined by the difference between the ion concentration in the reservoir connected to the base of the channel, $\rho_b$, and the concentration in the reservoir connected to the tip, $\rho_t$. Consequently, a diffusive fluid flow arises. However, we consider only the pressure and potential contributions to the total fluid flow, $Q = Q_P + Q_V$, assuming the diffusive flow to be negligible.  The resulting fluid flow generates non-trivial ion concentration profiles which, in turn, directly influence the channel's conductance $g$. 
%It has already been shown that, for $\Delta P = 0$, the channel shows current rectification, thus working as a diode \cite{}, due to concentration polarization.

@@

The PNN is composed of $\N$ nodes, an example is shown in \cref{fig:memristor_network}. 
\begin{figure}[t]
    \centering
    \includegraphics[width=0.6\columnwidth]{plots/memristor/memristor_network.pdf}
    \caption{An example of a network of five nodes connected by six memristors. The nodes are divided in input nodes (green circles), hidden nodes (grey nodes) and output nodes (purple circles) each of them is characterized by a triplet of physical quantities, representing the potential, pressure, and density. Th resulting stimuli on the memristor is given by the difference between base and tip signal, it thus depends on the orientation of the memristor. In the case of the memristor that connects the first and the second node, for example, the external stimuli are $\Delta V = V_2 - V_1$, $\Delta P = P_2 - P_1$ and $\Delta \rho = \rho_2 - \rho_1$. The zig-zag lines directed towards the input nodes indicate where the electrical signal is applied, while the zig-zag lines pointing outwards from the nodes indicate where the electrical signal is measured.}
    \label{fig:memristor_network}
\end{figure} 
Each node $\n\in[1,\N]$ is characterized by a value of electrical potential $V_{\n}$, pressure $P_{\n}$ and bulk density $\rho_{\n}$. Even though each node in the network is characterized by three physical properties, only electrical current flows through the network. This current propagates via the memristive elements, enabling modifications to the electric potential at other nodes. In contrast, pressure and density are externally imposed at each node and do not propagate through the network. The set of nodes is divided in three sets. A node belongs to the set of \textit{input nodes} if it is used to input current in the circuit, its value of voltage is known and fixed. After the inputs are imposed, the electrical signal travels through the network and the resulting output potential is read in the nodes belonging to \textit{output nodes}. The remaining nodes, that do not function as input or output, are included in the set of \textit{hidden nodes}.

The nodes in the PNN are connected by $\M$ memristive elements. Consequently, a memristor connecting two nodes experiences potential drop, pressure drop and concentration gradient given by the difference between the node where the base of the memristor is connected to, and the node to which the tip is linked. The memristor is characterized by its conductance state $g_{\m}$ with $\m\in[1,\M]$, consequence of both the geometry parameters and the external stimuli.
To solve numerically the resulting circuit, it is necessary to simultaneously address the circuit equations and the time evolution equation for the memristors' conductance. 
The circuit equations are handled using the open-source circuit solver \textsc{AHKAB}\cite{}, which computes the electric potentials at the nodes and the currents through the edges discretizing the differential equations in timesteps $\Delta t$. To incorporate the memristive elements, we approximate the time evolution of the conductances using a discretized form of \cref{eq:eqofmotconductance}, employing the Euler-forward method. Specifically, the conductance at a given time-step is calculated from its value at the previous time-step as:
\begin{multline}
    g_{\m}\brac{\boldsymbol{\Delta}(t), t} = g_{\m}\brac{\boldsymbol{\Delta}(t), t-\Delta t} + \\
    \frac{g_{\m,\infty}\brac{\boldsymbol{\Delta}(t)-g_{\m}\brac{\boldsymbol{\Delta}, t}}}{\tau}\Delta t.
\end{multline}
In this work, we only apply constant voltage supplies to the input nodes of the circuit. Nonetheless, we notice that solving the dynamical equations is needed due to the dynamical change in the conductance of the memristors that, in turn, results in time-dependent potential drops across them. A more detailed illustration of the algorithm that solves the circuit, along with an analysis of the relaxation to the steady state of the memristor network, is provided in \cref{app:dyn_evol_steady_state}.

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=1.7\columnwidth]{plots/voltage_divider/mse_weights_vd.pdf}
    \caption{(a) Behaviour of the cost function during the training of the memristor voltage divider. Different colors indicate that the training is done with a specific choice of training weights, as indicated in the legend. Learning rates are set to $[3\cdot 10^{-6}, 3\cdot 10^{-6}, 5\cdot 10^{-4}, 100]$. (b) Pressures imposed on the three nodes of the voltage divider during training. The inset shows the relaxation to the steady state of the output voltage $V_2$ to the desired value $V_2^D$ when the pressures imposed are the values obtained in the last training step, and the input nodes are stimulated with fixed voltage. (c) Ion concentration densities imposed on the three nodes of the voltage divider during training. (d) Length of the memristors during training. (e) Base radius of the memristors during training.}\label{fig:mse_weights_vd}
\end{figure*} 
In the forward phase, the network is fed with specific electric potentials in the input nodes. Following both circuit laws and memristor's dynamics, the network reaches a steady state where electrical potentials can be measured in the output nodes. Clearly, the output of the network depends on the set of weights imposed on the network $\boldsymbol{w}$, regardless them being node or edge defined. The value of the outputs can be used to quantify the status of the network, meaning how much its answer differs from the desired one. Specifically, we define a \emph{cost function} that computes the difference squared between the voltage in the output nodes and the values of desired voltages
\[
C\left( \boldsymbol{w} \right) = \sum_{i \in \Lambda} \brac{V_i\brac{\boldsymbol{w}} - V_i^D}^2,
\label{eq:cost_func}
\]
where $\Lambda$ represents the set of indices belonging to the output nodes, $V_i\brac{\boldsymbol{w}}$ the voltage measured on the output nodes and $V^D$ is a vector containing the values of desired output nodes.

@@

% \cleardoublepage
\subsection{\label{sec:linear_regression}Linear Regression}
\begin{figure*}[t]
    \centering
    \includegraphics[width=0.85\textwidth]{plots/regression/regression_G00010001.pdf}
    \caption{(a) Network geometry used for linear regression training. The network comprises $17$ edges connecting $10$ nodes, including one output node (purple dot) and three input nodes (green dots). Among the three input nodes, one is used to ground the circuit ($V_2 = 0V$), one ($V_4$) provides a constant voltage source for bias, and the remaining one ($V_4$) is used to input the training data point. (b) Test cost function evaluated every $10$ training steps for four different weight choices. Insets show visual representations of the network’s output trained with varying channel lengths at three different training steps.}\label{fig:lin_regr1}
\end{figure*}

@@

To evaluate how well the algorithm performs during training, we compute the \emph{test cost function} $ C_{\text{test}} $ over the entire test set. If $ \boldsymbol{w} $ denotes the weight vector at a given training step, the test cost is defined as
\[
C_{\text{test}}\left( \boldsymbol{w} \right) = \sum_{j=1}^{N_{\text{test}}} \left[ V_{\text{out}}(\boldsymbol{w}) - V_{\text{out}}^{(j)} \right]^2,
\]
which sums the squared differences between the measured and target output voltages at each test point. For each input $V_{\text{in}}^{(j)}$ in the test set, the voltage $ V_{\text{out}}(\boldsymbol{w}) $ is the output produced by the network with weights $ \boldsymbol{w} $, while $ V_{\text{out}}^{(j)} $ is the desired output. This test cost is used for performance evaluation and is not involved in the training process.

@@

The performance of the algorithm on this task is shown in \cref{fig:lin_regr2}(b), where the learning rates for the different weights have been fine-tuned to prevent overshooting while still maintaining comparable convergence speed. From the behaviour of the cost function, it is evident that the choice of weight type has a decisive impact on the network’s ability to learn a complex task. Specifically, training the lengths of the channels on the edges results in successful convergence, whereas training the node pressures proves to be less efficient. For a visual interpretation, insets in \cref{fig:lin_regr1}(b) show the state of the training at three different time steps when the selected weight is the length of the channel.
\begin{figure*}[t]
    \centering
    \includegraphics[width=0.85\textwidth]{plots/regression/regression_G00010002.pdf}
    \caption{(a) Network geometry used for linear regression training. It is equivalent to the geometry shown in \cref{fig:lin_regr1}, but the direction of the memristors highlighted in orange has been inverted. (b) Test cost function evaluated every $10$ training steps for four different weight choices. Insets show visual representations of the network’s output trained with varying node ion density at two different training steps.}\label{fig:lin_regr2}
\end{figure*}

@@

% \begin{algorithm}
%     \caption{Relaxation of a Memristor Voltage Divider to Steady State}
%     % \label{alg:relaxation}
%     \begin{algorithmic}[1]
%     \State \textbf{Input:} Initial conductance $g_0$, time step $\Delta t$, convergence threshold $\epsilon$, relaxation time $\tau$
%     \State Initialize $g_1 \gets g_0$, $g_2 \gets g_0$, $t \gets 0$
%     \While{not converged}
%         \State Compute $\Delta V_1 \gets \frac{5 \cdot g_2}{g_1 + g_2}$
%         \State Compute $\Delta V_2 \gets \frac{5 \cdot g_1}{g_1 + g_2}$
%         \State $g_1 \gets g_1 + \frac{g_{\infty}(\Delta V_1) - g_1}{\tau} \cdot \Delta t$
%         \State $g_2 \gets g_2 + \frac{g_{\infty}(\Delta V_2) - g_2}{\tau} \cdot \Delta t$
%         \If{$|g_1 - g_{\infty}(\Delta V_1)| < \epsilon$ \textbf{and} $|g_2 - g_{\infty}(\Delta V_2)| < \epsilon$}
%             \State \textbf{break}
%         \EndIf
%     \EndWhile
%     \end{algorithmic}
% \end{algorithm}

>>>>

\title{Multistimulus Machine Learning in Microfluidic Memristor Networks}

\author{Monica Conte}
\affiliation{Soft Condensed Matter \& Biophysics, Debye Institute for Nanomaterials Science, Utrecht University, Princetonplein 1, 3584 CC Utrecht, The Netherlands,}
\author{René van Roij}
\affiliation{Institute for Theoretical Physics, Utrecht University, Princetonplein 5, 3584 CC Utrecht, The Netherlands.}
\author{Marjolein Dijkstra}
\affiliation{Soft Condensed Matter \& Biophysics, Debye Institute for Nanomaterials Science, Utrecht University, Princetonplein 1, 3584 CC Utrecht, The Netherlands,}

@@

\begin{abstract}
Inspired by the brain's energy-efficient processing, we propose a physical implementation of artificial neural networks to surpass digital systems in thier capabilities. We consider a network with nodes connected by conical microfluidic memristor channels filled with aqueous electrolyte [1]. These memristive channels exhibit tunable conductance, in response to electrical, chemical, and mechanical stimuli, as well as to changes in channel geometry. Leveraging this multistimulus adaptability, we develop a hybrid training algorithm that treats either the channel geometry or external stimuli as adjustable weights. In this framework, the system itself performs the feedforward pass through physical relaxation, while a combined physical-numerical gradient descent executes the backward pass. The network performs basic learning tasks, such as input-output voltage mapping and linear regression with bias.
\end{abstract}
\maketitle

@@

Applications of PNNs have been explored across diverse fields, including optics \cite{optics1}, electronics \cite{electronics1}, photonics \cite{acoustic1} and mechanical systems \cite{mechanics1}, among others. Many existing training algorithms rely entirely on in silico methods, which often face challenges when transferring the learned model to real physical devices. To address this, hybrid training approaches have been proposed, where the inference phase is executed physically while the backward phase is computed numerically\cite{backporp_pnn}. Other methods, based on coupled learning\cite{supervised_learning}, implement local learning rules at the level of each connection or edge\cite{demonstr_coupled_learning}. Applying local rules removed the need of the optimiziation of a global cost function during learning, as in standard backpropagation methods. However, these approaches still face the need for two identical physical systems to apply different boundary conditions. In this work, we propose a novel algorithm in which both inference and backpropagation are carried out mainly taking advantage of the physical processes inherent in the system, specifically leveraging the natural relaxation dynamics of the system in response to external stimuli and weight modifications.

@@

The paper is organized as follows. \cref{sec:memristor_network} presents the physical network we define using memristor elements, called memristor network, and explains how this system is numerically solved. \cref{sec:training} describes the training algorithm that we use to achieve desired responses of the memristor network. In \cref{sec:train_memr_vd}, we exemplify the proposed algorithm by training a simple system, namely a voltage divider, for input-output voltage mapping. \cref{sec:train_memr_nw} shows a bigger network able to be trained for more complex input-output relationships. Finally, in \cref{sec:linear_regression} we adapt the training algorthm to train a network to perform a linear regression with bias. This section is followed by conclusions and suggestions for future work in \cref{sec:concl}.
% \newpage

@@

At the far side of both reservoirs, electrical potential, pressure and ion concentration are imposed. As a result, the memristor feels a potential drop $\Delta V = V_b - V_t$ defined as the difference between the potential in the base and in the tip. The ions within the EDL region generate an electro-osmotic fluid flow $Q_V = (\pi R_b R_t\epsilon\psi_0/L\eta)\Delta V$. Meanwhile, the pressure drop $\Delta P = P_b - P_t$, given by the difference between the base and the tip pressure, induces a Poiseuille-like flow $Q_P = (\pi R_b^3 R_t^3/8L\eta\langle R^2\rangle)\Delta P$. Finally, the ions experience a salt concentration gradient $\Delta \rho = \rho_b - \rho_t$, defined by the difference between the ion concentration in the reservoir connected to the base of the channel, $\rho_b$, and the concentration in the reservoir connected to the tip, $\rho_t$. Consequently, a diffusive fluid flow arises. However, we consider only the pressure and potential contributions to the total fluid flow, $Q = Q_P + Q_V$, assuming the diffusive flow to be negligible.  The resulting fluid flow generates non-trivial ion concentration profiles which, in turn, directly influence the channel's conductance $g$. 
%It has already been shown that, for $\Delta P = 0$, the channel shows current rectification, thus working as a diode \cite{}, due to concentration polarization.

@@

The PNN is composed of $N$ nodes, an example is shown in \cref{fig:memristor_network}. 
\begin{figure}[t]
    \centering
    \includegraphics[width=0.6\columnwidth]{plots/memristor/memristor_network.pdf}
    \caption{An example of a network of five nodes connected by six memristors. The nodes are divided in input nodes (green circles), hidden nodes (grey nodes) and output nodes (purple circles) each of them is characterized by a triplet of physical quantities, representing the potential, pressure, and density. Th resulting stimuli on the memristor is given by the difference between base and tip signal, it thus depends on the orientation of the memristor. In the case of the memristor that connects the first and the second node, for example, the external stimuli are $\Delta V = V_2 - V_1$, $\Delta P = P_2 - P_1$ and $\Delta \rho = \rho_2 - \rho_1$. The zig-zag lines directed towards the input nodes indicate where the electrical signal is applied, while the zig-zag lines pointing outwards from the nodes indicate where the electrical signal is measured.}
    \label{fig:memristor_network}
\end{figure} 
Each node $\n\in[1,N]$ is characterized by a value of electrical potential $V_{\n}$, pressure $P_{\n}$ and bulk density $\rho_{\n}$. Even though each node in the network is characterized by three physical properties, only electrical current flows through the network. This current propagates via the memristive elements, enabling modifications to the electric potential at other nodes. In contrast, pressure and density are externally imposed at each node and do not propagate through the network. The set of nodes is divided in three sets. A node belongs to the set of \textit{input nodes} if it is used to input current in the circuit, its value of voltage is known and fixed. After the inputs are imposed, the electrical signal travels through the network and the resulting output potential is read in the nodes belonging to \textit{output nodes}. The remaining nodes, that do not function as input or output, are included in the set of \textit{hidden nodes}.

The nodes in the PNN are connected by $M$ memristive elements. Consequently, a memristor connecting two nodes experiences a potential drop, pressure drop and concentration gradient given by the difference between the node where the base of the memristor is connected to, and the node to which the tip is linked. The memristor is characterized by its conductance state $g_{\m}$ with $\m\in[1,M]$, consequence of both the geometry parameters and the external stimuli.
To solve numerically the resulting circuit, it is necessary to simultaneously address the circuit equations and the time evolution equation for the memristors' conductance. 
The circuit equations are handled using the open-source circuit solver \textsc{AHKAB}\cite{}, which computes the electric potentials at the nodes and the currents through the edges discretizing the differential equations in timesteps $\Delta t$. To incorporate the memristive elements, we approximate the time evolution of the conductances using a discretized form of \cref{eq:eqofmotconductance}, employing the Euler-forward method. Specifically, the conductance at a given time-step is calculated from its value at the previous time-step as:
\begin{multline}
    g_{\m}\brac{\boldsymbol{\Delta}(t), t} = g_{\m}\brac{\boldsymbol{\Delta}(t), t-\Delta t} + \\
    \frac{g_{\m,\infty}\brac{\boldsymbol{\Delta}(t)-g_{\m}\brac{\boldsymbol{\Delta}, t}}}{\tau}\Delta t.
\end{multline}
In this work, we only apply constant voltage supplies to the input nodes of the circuit. Nonetheless, we notice that solving the dynamical equations is needed due to the dynamical change in the conductance of the memristors that, in turn, results in time-dependent potential drops across them. A more detailed illustration of the algorithm that solves the circuit, along with an analysis of the relaxation to the steady state of the memristor network, is provided in \cref{app:dyn_evol_steady_state}.

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=1.7\columnwidth]{plots/voltage_divider/mse_weights_vd.pdf}
    \caption{(a) Behaviour of the cost function $C(\boldsymbol{w})$ during the training of the memristor voltage divider. Different colors indicate that the training is done with a specific choice of training weights, as indicated in the legend. Learning rates are set to $[3\cdot 10^{-6}, 3\cdot 10^{-6}, 5\cdot 10^{-4}, 100]$. (b) Pressures imposed on the three nodes of the voltage divider during training. The inset shows the relaxation to the steady state of the output voltage $V_2$ to the desired value $V_2^D$ when the pressures imposed are the values obtained in the last training step, and the input nodes are stimulated with fixed voltage. (c) Ion concentration densities imposed on the three nodes of the voltage divider during training. (d) Length of the memristors during training. (e) Base radius of the memristors during training.}\label{fig:mse_weights_vd}
\end{figure*} 
In the forward phase, the network is fed with specific electric potentials in the input nodes. Following both circuit laws and memristor's dynamics, the network reaches a steady state where electrical potentials can be measured in the output nodes. Clearly, the output of the network depends on the set of weights imposed on the network $\boldsymbol{w}$, regardless them being node or edge defined. The value of the outputs can be used to quantify the status of the network, meaning how much its answer differs from the desired one. Specifically, we define a \emph{cost function} that computes the difference squared between the voltage in the output nodes and the values of desired voltages
\[
C\left( \boldsymbol{w} \right) = \sum_{i \in \Lambda} \brac{V_i\brac{\boldsymbol{w}} - V_i^D}^2,
\label{eq:cost_func}
\]
where $\Lambda$ represents the set of indices belonging to the output nodes, $V_i\brac{\boldsymbol{w}}$ the voltage measured on the output nodes and $V^D$ is a vector containing the values of desired output nodes.

@@

\subsection{\label{sec:linear_regression}Linear Regression}
\begin{figure*}[t]
    \centering
    \includegraphics[width=0.85\textwidth]{plots/regression/regression_G00010001.pdf}
    \caption{(a) Network geometry used for linear regression training. The network comprises $17$ edges connecting $10$ nodes, including one output node (purple dot) and three input nodes (green dots). Among the three input nodes, one is used to ground the circuit ($V_2 = 0V$), one ($V_4$) provides a constant voltage source for bias, and the remaining one ($V_4$) is used to input the training data point. For the four weight types (channle length, base radius, ion density and pressure), we used as constant voltage input $V_4 = [11V,4V,4V,11V]$ and learning rates $[2\cdot 10^{-7}, 1\cdot 10^{-6}, 1\cdot 10^{-4}, 2\cdot 10^{2}]$. (b) Test cost function evaluated every $10$ training steps for four different weight choices. Insets show visual representations of the network’s output trained with varying channel lengths at three different training steps.}\label{fig:lin_regr1}
\end{figure*}

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=0.85\textwidth]{plots/regression/regression_G00010002.pdf}
    \caption{(a) Network geometry used for linear regression training. It is equivalent to the geometry shown in \cref{fig:lin_regr1}, but the direction of the memristors highlighted in orange has been inverted. For the four weight types (channle length, base radius, ion density and pressure), we used as constant voltage input $V_4 = [11V,4V,4V,11V]$ and learning rates $[3\cdot 10^{-7}, 1\cdot 10^{-6}, 9\cdot 10^{-4}, 2\cdot 10^{2}]$. (b) Test cost function evaluated every $10$ training steps for four different weight choices. Insets show visual representations of the network’s output trained with varying node ion density at two different training steps.}\label{fig:lin_regr2}
\end{figure*} 
To evaluate how well the algorithm performs during training, we compute the \emph{test cost function} $ C_{\text{test}} $ over the entire test set. If $ \boldsymbol{w} $ denotes the weight vector at a given training step, the test cost is defined as
\[
C_{\text{test}}\left( \boldsymbol{w} \right) = \sum_{j=1}^{N_{\text{test}}} \left[ V_{\text{out}}(\boldsymbol{w}) - V_{\text{out}}^{(j)} \right]^2,
\]
which sums the squared differences between the measured and target output voltages at each test point. For each input $V_{\text{in}}^{(j)}$ in the test set, the voltage $ V_{\text{out}}(\boldsymbol{w}) $ is the output produced by the network with weights $ \boldsymbol{w} $, while $ V_{\text{out}}^{(j)} $ is the desired output. This test cost is used for performance evaluation and is not involved in the training process.

@@

The performance of the algorithm on this task is shown in \cref{fig:lin_regr2}(b), where the learning rates for the different weights have been fine-tuned to prevent overshooting while still maintaining comparable convergence speed. From the behaviour of the cost function, it is evident that the choice of weight type has a decisive impact on the network’s ability to learn a complex task. Specifically, training the lengths of the channels on the edges results in successful convergence, whereas training the node pressures proves to be less efficient. For a visual interpretation, insets in \cref{fig:lin_regr1}(b) show the state of the training at three different time steps when the selected weight is the length of the channel.

@@

\begin{algorithm}
    \caption{}
    \label{alg:relaxation}
    \begin{algorithmic}[1]
    \State \textbf{Input:} Initial conductance $g_0$, time step $\Delta t$, convergence threshold $\epsilon$, relaxation time $\tau$
    \State Initialize $g_1 \gets g_0$, $g_2 \gets g_0$, $t \gets 0$
    \While{not converged}
        \State Compute $\Delta V_1 \gets \frac{5 \cdot g_2}{g_1 + g_2}$
        \State Compute $\Delta V_2 \gets \frac{5 \cdot g_1}{g_1 + g_2}$
        \State $g_1 \gets g_1 + \frac{g_{\infty}(\Delta V_1) - g_1}{\tau} \cdot \Delta t$
        \State $g_2 \gets g_2 + \frac{g_{\infty}(\Delta V_2) - g_2}{\tau} \cdot \Delta t$
        \If{$|g_1 - g_{\infty}(\Delta V_1)| < \epsilon$ \textbf{and} $|g_2 - g_{\infty}(\Delta V_2)| < \epsilon$}
            \State \textbf{break}
        \EndIf
    \EndWhile
    \end{algorithmic}
\end{algorithm}
